# Scenario Analysis: WCS Priority Landscapes

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls())
gc()
```

```{r, include=FALSE, eval=TRUE, echo=FALSE, warning=FALSE}
# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  , results='hide'
  , fig.width = 10
  , fig.height = 8
)
# bread-and-butter
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)
library(latex2exp)
# visualization
library(kableExtra)
library(cowplot)
library(RColorBrewer)
library(ggtext)
library(mapview) #Interactive maps
library(leafpop) #map html popup
library(tidytext) # reorder within facets
# spatial analysis
library(sf)
library(USAboundaries)
library(ggmap) # nice ggplot maps
# set seed
set.seed(11)
# crs
transform_crs <- 5070
```

```{r, results='hide'}
# turn off the s2 processing 
## https://stackoverflow.com/questions/68478179/how-to-resolve-spherical-geometry-failures-when-joining-spatial-data
sf::sf_use_s2(FALSE)
```

See the USFS [Wildfire Crisis Strategy]((https://www.fs.usda.gov/managing-land/wildfire-crisis)) (WCS) document for background information.

## Data Preparation

### Load Disneyland

```{r}
# point
pt <- data.frame(
  lat = c(38.89373169777448)
  , lon = c(-77.03675843174503)
)
dc <- st_as_sf(pt, coords = c("lon","lat"), remove = FALSE, crs = "EPSG:4326") %>% 
  sf::st_transform(transform_crs)
  sf::st_buffer(1000)



### disney spatial data
disney <- sf::st_read("dis_disneyland_pc.shp") %>%
  sf::st_transform(transform_crs) 
  # sf::st_centroid() %>% 
  # sf::st_buffer(1000)
  
# ### disney spatial data
# disney2 <- sf::st_read("dis_adventureland_pc.shp") %>%
#   sf::st_transform(transform_crs)
# ggplot() +
#   geom_sf(data = disney, fill = NA, color = "navy") +
#   geom_sf(data = disney2, fill = NA, color = "gray") +
#   theme_void()

mapview::mapviewOptions(homebutton = FALSE, basemaps = c("OpenStreetMap","Esri.WorldImagery"))
mapview::mapview(dc
        , color = "black"
        , lwd = 3
        , alpha.regions = 0
        , label = FALSE
        , legend = FALSE
        , popup = FALSE
        # , layer.name = "Disneyland"
        , layer.name = "DC"
  )
```

### Load fireshed project area data

The fireshed registry spatial data was obtained from the USFS Geospatial Data Discovery tool for [firesheds](https://data-usfs.hub.arcgis.com/datasets/fireshed-registry-fireshed-feature-layer/explore) and [project areas](https://data-usfs.hub.arcgis.com/datasets/fireshed-registry-project-area-feature-layer/explore) (accessed 2023-05-03).

See [this section](#fsheds) for exploration of fireshed and project area spatial data

```{r fshed-spatial}
### fireshed spatial data
fireshed <- sf::st_read("C:/Data/usfs/wcs_mechanical_constraints/data/firesheds/Fireshed_Registry3A_Fireshed/Fireshed_Registry%3A_Fireshed_(Feature_Layer).shp") %>%
  sf::st_transform(transform_crs) %>% 
  setNames(c(
      "shape_id"
      , "area_ha"
      , "fireshed_id"
      , "fireshed_name"
      , "fireshed_code"
      , "fireshed_state"
      , "nopas"
      , "objectid"
      , "fshed_id"
      , "exp_total"
      , "exp_usfs"
      , "exp_nonfs"
      , "exp_usfs_protected"
      , "exp_nonfs_protected"
      , "exp_usfs_managed"
      , "exp_nonfs_managed"
      , "exp_usfs_forest"
      , "exp_nonfs_forest"
      , "exp_usfs_nonforest"
      , "exp_nonfs_nonforest"
      , "exp_usfs_conifer"
      , "exp_nonfs_conifer"
      , "exp_usfs_managedforest"
      , "exp_nonfs_managedforest"
      , "exp_usfs_managedconifer"
      , "exp_nonfs_managedconifer"
      , "exp_nonfs_nonconifer_hihaz"
      , "dist_vs"
      , "crisis_strategy"
      , "key_preformance_indicator"
      , "national_usfs_rank"
      , "national_all_land_rank"
      , "regional_usfs_rank"
      , "regional_all_land_rank"
      , "start_date"
      , "end_date"
      , "geometry"
  )) %>% 
  dplyr::mutate(
    exposure_pct_rank = dplyr::percent_rank(exp_total)
    , exposure_pct_rank_grp = dplyr::case_when(
      exposure_pct_rank >= 1-0.01 ~ "Top 1%"
      , exposure_pct_rank >= 1-0.05 ~ "Top 5%"
      , exposure_pct_rank >= 1-0.10 ~ "Top 10%"
      , exposure_pct_rank >= 1-0.25 ~ "Top 25%"
      , TRUE ~ "Bottom 75%"
    ) %>% 
    factor(
      levels = c("Top 1%","Top 5%","Top 10%","Top 25%","Bottom 75%")
      , ordered = T
    )
    # there is also a national_all_land_rank column
    , ntllandrank_pct_rank = dplyr::percent_rank(-national_all_land_rank)
    , ntllandrank_pct_rank_grp = dplyr::case_when(
        ntllandrank_pct_rank >= 1-0.01 ~ "Top 1%"
        , ntllandrank_pct_rank >= 1-0.05 ~ "Top 5%"
        , ntllandrank_pct_rank >= 1-0.10 ~ "Top 10%"
        , ntllandrank_pct_rank >= 1-0.25 ~ "Top 25%"
        , TRUE ~ "Bottom 75%"
      ) %>% 
      factor(
        levels = c("Top 1%","Top 5%","Top 10%","Top 25%","Bottom 75%")
        , ordered = T
      )
    , crisis_strategy = ifelse(is.na(crisis_strategy),"Not High Risk",crisis_strategy) %>% 
      as.factor() %>% 
      forcats::fct_shift()
  )
  #rename sf geom column
    names(fireshed)[names(fireshed)==tolower(attr(fireshed, "sf_column"))] = "geometry"
    sf::st_geometry(fireshed) = "geometry"
    # calculate area
    fireshed <- fireshed %>% 
      dplyr::mutate(
        fireshed_area_ha = as.numeric(sf::st_area(geometry))/10000
        , fireshed_area_acres = (fireshed_area_ha*10000)/4046.85642
      )
## fireshed_proj_area spatial data
fireshed_proj_area <- sf::st_read("C:/Data/usfs/wcs_mechanical_constraints/data/firesheds/Fireshed_Registry3A_Project_Area/Fireshed_Registry%3A_Project_Area_(Feature_Layer).shp") %>%
  sf::st_transform(transform_crs) %>% 
  setNames(c(
      "shape_id"
      , "fireshed_id"
      , "pa_id"
      , "pa_area_ha"
      , "objectid"
      , "pa_id2"
      , "fshed_id"
      , "exp_total"
      , "exp_usfs"
      , "exp_nonfs"
      , "exp_usfs_protected"
      , "exp_nonfs_protected"
      , "exp_usfs_managed"
      , "exp_nonfs_managed"
      , "exp_usfs_forest"
      , "exp_nonfs_forest"
      , "exp_usfs_nonforest"
      , "exp_nonfs_nonforest"
      , "exp_usfs_conifer"
      , "exp_nonfs_conifer"
      , "exp_usfs_managedforest"
      , "exp_nonfs_managedforest"
      , "exp_usfs_managedconifer"
      , "exp_nonfs_managedconifer"
      , "exp_nonfs_nonconifer_hihaz"
      , "dist_vs"
      , "pctrecentlydisturbed"
      , "start_date"
      , "end_date"
      , "geometry"
  )) %>% 
  dplyr::mutate(
    exposure_pct_rank = dplyr::percent_rank(exp_total)
    , exposure_pct_rank_grp = dplyr::case_when(
      exposure_pct_rank >= 1-0.01 ~ "Top 1%"
      , exposure_pct_rank >= 1-0.05 ~ "Top 5%"
      , exposure_pct_rank >= 1-0.10 ~ "Top 10%"
      , exposure_pct_rank >= 1-0.25 ~ "Top 25%"
      , TRUE ~ "Bottom 75%"
    ) %>% 
    factor(
      levels = c("Top 1%","Top 5%","Top 10%","Top 25%","Bottom 75%")
      , ordered = T
    )
  )
  #rename sf geom column
    names(fireshed_proj_area)[names(fireshed_proj_area)==tolower(attr(fireshed_proj_area, "sf_column"))] = "geometry"
    sf::st_geometry(fireshed_proj_area) = "geometry"
    # calculate area
    fireshed_proj_area <- fireshed_proj_area %>% 
      dplyr::mutate(
        pa_area_ha = as.numeric(sf::st_area(geometry))/10000
        , pa_area_acres = (pa_area_ha*10000)/4046.85642
      ) %>% 
      # JOIN WITH FIRESHED DATA
      dplyr::inner_join(
        fireshed %>%
          sf::st_drop_geometry() %>%
          dplyr::select(fireshed_id, crisis_strategy, exp_total
                        , exposure_pct_rank, exposure_pct_rank_grp
          ) %>% 
          dplyr::rename(exposure_total=exp_total) %>% 
          dplyr::rename_with(
            ~ paste0("fireshed_",.x)
            , -c(fireshed_id)
          )
        , by = dplyr::join_by(fireshed_id)
      ) %>%
      dplyr::select(pa_id,pa_area_ha
                    ,exp_total,exposure_pct_rank,exposure_pct_rank_grp
                    , tidyselect::starts_with("fireshed_")
      ) %>% 
      dplyr::rename(exposure_total=exp_total) %>%
      dplyr::rename_with(
        ~ paste0("pa_", .x, recycle0 = TRUE)
        , tidyselect::starts_with("exp")
      )
```

## Combine and plot

```{r}
fshed_disney <- fireshed %>% 
  sf::st_filter(disney)
fshed_pa_disney <- fireshed_proj_area %>% 
  dplyr::filter(fireshed_id %in% fshed_disney$fireshed_id)


sf::st_write(fshed_disney %>% 
               dplyr::select(fireshed_id, fireshed_name)
               ,"fshed_disney.shp", append = F)
sf::st_write(fshed_pa_disney %>%
               dplyr::select(fireshed_id, pa_id)
             ,"fshed_pa_disney.shp", append = F)

mapview(fshed_pa_disney
        , color = "navy"
        , lwd = 1
        , alpha.regions = 0
        , label = FALSE
        , legend = FALSE
        , popup = FALSE
        , layer.name = "fpa"
        ) +
mapview::mapview(disney
        , color = "black"
        , lwd = 3
        , alpha.regions = 0
        , label = FALSE
        , legend = FALSE
        , popup = FALSE
        , layer.name = "Disneyland"
  ) 
  
```

upload .shp files to GEE and [run this script](https://code.earthengine.google.com/aee806da7765bd36dd81fcc9dcdf9036?noload=true) to export Disneyland NAIP tif file.
